---
title: "LASSO Model for CMI-PB"
output: html_document
date: "2023-07-20"
---

### Load required libraries
```{r}
suppressWarnings({library(omicade4)
library(mogsa)
library(RSpectra)
library(lubridate)
library(dplyr)
library(cowplot)
library(ggplot2)
library(reshape2)
library(DBI)
# library(RPostgreSQL)
library(tidyr)
library(readr)
library(tibble)
library(tidyverse)
library(BatchQC)
library(sva)
library(RcmdrMisc)
library('rstatix')
library("Hmisc")
library(corrplot)
library(GGally)
library(gridExtra)
library(ggpubr)
library(data.table)
library(glmnet)
library(here)
  })
```


### Set up new directories
```{r}
DEMOGRAPHY_COLS <- c("age_at_boost", "infancy_vac", "biological_sex")

ABTITER_COLS <- c("IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0',
                   "IgG_PT_day14",'IgG_FHA_day14','IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14', 'IgG4_PT_day14','IgG4_FHA_day14')

CELL_COLS <- c("Monocytes_day0", "Monocytes_day1", "CD4Tcells_day0", 
               "CD4Tcells_day3", "ASCs..Plasmablasts._day0", "ASCs..Plasmablasts._day7")

RNA_COLS <- c("CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0", 
              "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14")


"Dataset 2020 ****************************************************************************************************************"
# subDir <- "dataset2020"
# dataFile <- "dataset2020.csv"
# readyToUseFile <- "dataset2020_logTransformed_encoded.csv"
# 
# # REMOVE_OUTLIERS <- TRUE
# REMOVE_OUTLIERS <- FALSE
# 
# LOG_TRANSFORM <- TRUE
# # LOG_TRANSFORM <- FALSE
# 
# LOG_TRANS_COLS <- c( "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
#                     'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')
# 
# # NEGATIVE_COLS <-RNA_COLS
# NEGATIVE_COLS <- c()

"Dataset 2021 ****************************************************************************************************************"
subDir <- "dataset2021"
dataFile <- "dataset2021.csv"
readyToUseFile <- "dataset2021_logTransformed_encoded.csv"

# REMOVE_OUTLIERS <- TRUE
REMOVE_OUTLIERS <- FALSE

LOG_TRANSFORM <- TRUE
# LOG_TRANSFORM <- FALSE

LOG_TRANS_COLS <- c("CCL3_day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0',
                    'IgG4_PT_day0','IgG4_FHA_day0', "IL6_day0", "NFKBIA_day0", "XIST_day0",
                    "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
                    'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')

# NEGATIVE_COLS <-RNA_COLS
NEGATIVE_COLS <- c()

"Combined Dataset ****************************************************************************************************************"
# subDir <- "combined_datasets2020_2021"
# dataFile <- "combined_dataset2020_2021_batchCorrected.csv"
# readyToUseFile <- "combined_dataset2020_2021_batchCorrected_logTransformed_encoded.csv"
# 
# # REMOVE_OUTLIERS <- TRUE
# REMOVE_OUTLIERS <- FALSE
# 
# LOG_TRANSFORM <- TRUE
# # LOG_TRANSFORM <- FALSE
# 
# LOG_TRANS_COLS <- c("CCL3_day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0',
#                     'IgG4_PT_day0','IgG4_FHA_day0', "IL6_day0", "NFKBIA_day0", "XIST_day0",
#                     "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
#                     'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')
# 
# NEGATIVE_COLS <-RNA_COLS
# # NEGATIVE_COLS <- c()

"Create subdirectories ****************************************************************************************************************"
dir.create(here(file.path('./results', subDir)), showWarnings = FALSE)

subDir_multiPredictor <- "3_multiPredictor"
dir.create(here(file.path('./results', subDir, subDir_multiPredictor)), showWarnings = FALSE)
dir.create(here(file.path('./results', subDir, "3_multiPredictor/LASSO")), showWarnings = FALSE)
```

### Function for regression plot
```{r}
regPlot <- function(df, xCol, yCol,title){
  df$infancy_vac[df$infancy_vac==1] <- "wP"
  df$infancy_vac[df$infancy_vac==2] <- "aP"
  
  p <- ggplot(data = df, aes_string(x= xCol, y = yCol, label= "subject_id")) +
  geom_text(aes(colour=factor(infancy_vac))) +
  scale_color_manual(values=c("red", "springGreen4")) +
  labs(colour= "Vaccine Type") +
  stat_smooth(method = "lm", col = "dodgerblue3") +
  theme(panel.background = element_rect(fill = "white"),
  axis.line.x=element_line(),
  axis.line.y=element_line()) +
  ggtitle(title)+
  stat_cor(method = "spearman")
  }
```


### 3) LASSO model with multiple predictors
Baseline of all three tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

1. only demographic
2. only 3 baseline 
3. 3 baseline + demographic
4. 14 baseline
5. 14 baseline + demographic


### 1. only demographic
## leave one out cross validation been used to estimate model quality
```{r}
df <- read.csv(here(paste("./results/", subDir, "/", readyToUseFile, sep="")))
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

pval_pred_true <- data.frame(row.names = y)
correlation_pred_true <- data.frame(row.names = y)


for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  print(c(y[j], dim(subDf)))
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()

  for (i in 1:nrow(subDf)){
    oneout <- subDf[-c(i), ]
    
    suppressWarnings(fit_1<-cv.glmnet(x=as.matrix(oneout[, x]), oneout[,y[j]], 
                                          family='gaussian', alpha=1, nfolds=nrow(oneout)-1))
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    preds<-predict(fit_1,newx=as.matrix(predictor),s='lambda.min')
    predVals <- c(predVals, preds)
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  
  pval_pred_true[y[j], "Demographic"] <- as.matrix(correlations$p.value)
  correlation_pred_true[y[j], "Demographic"] <- as.matrix(correlations$estimate)
  
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  # # 
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography"))
  ggsave(p, file=here(paste('./results/', subDir, "/3_multiPredictor/LASSO/predictedTrue_multiPredictor_demographic_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}

```

## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


coefDf_multi <- data.frame(row.names = c(x, "(Intercept)"))
for (i in 1:length(y)){
  subDf <- na.omit(df[, c(x, y[i])])
  suppressWarnings(fit_1 <- cv.glmnet(x=as.matrix(subDf[,x]), subDf[,y[i]], 
                                      family='gaussian', alpha=1, nfolds=nrow(subDf)-1))
  coefDf_multi[y[i]] <- as.matrix(coef(fit_1, s = 'lambda.min'))
}

png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_regression_demography.png", sep="")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(as.matrix(t(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, number.cex = 1,
         col.lim =c(min(as.matrix(coefDf_multi), na.rm = TRUE), max(as.matrix(coefDf_multi), na.rm = TRUE)),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()


```

### 2. only baseline
## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CCL3_day0", "IgG_PT_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  print(c(y[j], dim(subDf)))
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()

  for (i in 1:nrow(subDf)){
    oneout <- subDf[-c(i), ]
    
    suppressWarnings(fit_1<-cv.glmnet(x=as.matrix(oneout[, x]), oneout[,y[j]], 
                                          family='gaussian', alpha=1, nfolds=nrow(oneout)-1))
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    preds<-predict(fit_1,newx=as.matrix(predictor),s='lambda.min')
    predVals <- c(predVals, preds)
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  
  pval_pred_true[y[j], "3 Tasks Baselines"] <- correlations$p.value
  correlation_pred_true[y[j], "3 Tasks Baselines"] <- correlations$estimate
  # # 
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  
  # #
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against 3 Baseline"))
  ggsave(p, file=here(paste('./results/', subDir, "/3_multiPredictor/LASSO/predictedTrue_multiPredictor_3baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}


```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CCL3_day0", "IgG_PT_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

coefDf_multi <- data.frame(row.names = c(x, "(Intercept)"))
for (i in 1:length(y)){
  subDf <- na.omit(df[, c(x, y[i])])
  suppressWarnings(fit_1 <- cv.glmnet(x=as.matrix(subDf[, x]), subDf[,y[i]], 
                                      family='gaussian', alpha=1, nfolds=nrow(subDf)-1))
  coefDf_multi[y[i]] <- as.matrix(coef(fit_1, s = 'lambda.min'))
}


png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_regression_3baseline.png", sep="")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, number.cex = 0.8,
         col.lim =c(min(as.matrix(coefDf_multi), na.rm = TRUE), max(as.matrix(coefDf_multi), na.rm = TRUE)),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```

### 3. Baseline of all three tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex", "Monocytes_day0", "IgG_PT_day0", "CCL3_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  print(c(y[j], dim(subDf)))
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()

  for (i in 1:nrow(subDf)){
    oneout <- subDf[-c(i), ]
    
    suppressWarnings(fit_1<-cv.glmnet(x=as.matrix(oneout[, x]), oneout[,y[j]], 
                                          family='gaussian', alpha=1, nfolds=nrow(oneout)-1))
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    preds<-predict(fit_1,newx=as.matrix(predictor),s='lambda.min')
    predVals <- c(predVals, preds)
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  
  pval_pred_true[y[j], "Demographic + 3 Baselines"] <- as.matrix(correlations$p.value)
  correlation_pred_true[y[j], "Demographic + 3 Baselines"] <- as.matrix(correlations$estimate)

  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType

  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography + 3 Baseline"))
  ggsave(p, file=here(paste('./results/', subDir, "/3_multiPredictor/LASSO/predictedTrue_multiPredictor_3baseline_demography_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}


```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex", "Monocytes_day0", "IgG_PT_day0", "CCL3_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

coefDf_multi <- data.frame(row.names = c(x, "(Intercept)"))
for (i in 1:length(y)){
  subDf <- na.omit(df[, c(x, y[i])])
  suppressWarnings(fit_1 <- cv.glmnet(x=as.matrix(subDf[, x]), subDf[,y[i]], 
                                      family='gaussian', alpha=1, nfolds=nrow(subDf)-1))
  coefDf_multi[y[i]] <- as.matrix(coef(fit_1, s = 'lambda.min'))
}


png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_regression_3baseline_demography.png", sep="")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, number.cex = 1,
         col.lim =c(min(as.matrix(coefDf_multi), na.rm = TRUE), max(as.matrix(coefDf_multi), na.rm = TRUE)),
          na.label=" ", tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```

### 4. 14 baselines
## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0',
       'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

corr_pred_true <- data.frame()

for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  print(c(y[j], dim(subDf)))
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()

  for (i in 1:nrow(subDf)){
    oneout <- subDf[-c(i), ]
    
    suppressWarnings(fit_1<-cv.glmnet(x=as.matrix(oneout[, x]), oneout[,y[j]], 
                                          family='gaussian', alpha=1, nfolds=nrow(oneout)-1))
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    preds<-predict(fit_1,newx=as.matrix(predictor),s='lambda.min')
    predVals <- c(predVals, preds)
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  
  pval_pred_true[y[j], "14 Tasks Baselines"] <- as.matrix(correlations$p.value)
  correlation_pred_true[y[j], "14 Tasks Baselines"] <- as.matrix(correlations$estimate)
  # # 
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType

  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against 14 Baselines"))
  ggsave(p, file=here(paste('./results/', subDir, "/3_multiPredictor/LASSO/predictedTrue_multiPredictor_14baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}

```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0',
       'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


coefDf_multi <- data.frame(row.names = c(x, "(Intercept)"))
for (i in 1:length(y)){
  subDf <- na.omit(df[, c(x, y[i])])
  suppressWarnings(fit_1 <- cv.glmnet(x=as.matrix(subDf[, x]), subDf[,y[i]], 
                                      family='gaussian', alpha=1, nfolds=nrow(subDf)-1))
  coefDf_multi[y[i]] <- as.matrix(coef(fit_1, s = 'lambda.min'))
}


png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_regression_14baseline.png", sep="")), width = 15, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,number.cex = 1.5,
         col.lim =c(min(as.matrix(coefDf_multi), na.rm = TRUE), max(as.matrix(coefDf_multi), na.rm = TRUE)),
          na.label=" ", tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue",  "white","red"))(100))
dev.off()

```

### 5. Baseline of all 14 tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("age_at_boost", "biological_sex", "infancy_vac", "Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", 
       "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0', 'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 
       'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  print(c(y[j], dim(subDf)))
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()

  for (i in 1:nrow(subDf)){
    oneout <- subDf[-c(i), ]
    
    suppressWarnings(fit_1<-cv.glmnet(x=as.matrix(oneout[, x]), oneout[,y[j]], 
                                          family='gaussian', alpha=1, nfolds=nrow(oneout)-1))
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    preds<-predict(fit_1,newx=as.matrix(predictor),s='lambda.min')
    predVals <- c(predVals, preds)
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  
  pval_pred_true[y[j], "Demographic + 14 Baselines"] <- as.matrix(correlations$p.value)
  correlation_pred_true[y[j], "Demographic + 14 Baselines"] <- as.matrix(correlations$estimate)
  
  # # 
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType

  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography + 14 Baseline"))
  ggsave(p, file=here(paste('./results/', subDir, "/3_multiPredictor/LASSO/predictedTrue_multiPredictor_14baseline_demographic", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}


png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_pred_true_spearman_0.95Confidence.png", sep="")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.05, number.cex = 1.5, tl.srt = 45, na.label=" ",
         cl.ratio = 0.2, col = colorRampPalette(c("blue",  "white","red"))(100))
dev.off()

png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_pred_true_spearman_0.99Confidence.png", sep="")), width = 10, height = 5  , units = 'in', res = 300 )
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.01, number.cex = 1.5, tl.srt = 45, na.label=" ",
         cl.ratio = 0.2, col = colorRampPalette(c("blue",  "white","red"))(100))
dev.off()

write.csv(pval_pred_true, here(paste('./results/', subDir, "/3_multiPredictor/LASSO/p_value_pred_true_spearman.csv", sep = "")))
write.csv(correlation_pred_true, here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlation_pred_true_spearman.csv", sep = "")))

```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "biological_sex", "infancy_vac", "Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", 
       "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0', 'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 
       'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14") # Monocyte has less values than samples


coefDf_multi <- data.frame(row.names = c(x, "(Intercept)"))
for (i in 1:length(y)){
  subDf <- na.omit(df[, c(x, y[i])])
  suppressWarnings(fit_1 <- cv.glmnet(x=as.matrix(subDf[, x]), subDf[,y[i]], 
                                      family='gaussian', alpha=1, nfolds=nrow(subDf)-1))
  coefDf_multi[y[i]] <- as.matrix(coef(fit_1, s = 'lambda.min'))
}


png(here(paste('./results/', subDir, "/3_multiPredictor/LASSO/correlatinHeatmap_regression_14baseline_demography.png", sep="")), width = 15, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,number.cex = 1.5,
         col.lim =c(min(as.matrix(coefDf_multi), na.rm = TRUE), max(as.matrix(coefDf_multi), na.rm = TRUE)), 
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue",  "white","red"))(100))
dev.off()

```
