---
title: "LM Model for CMIPB"
output: html_document
date: "2023-07-20"
"https://www.dataquest.io/blog/statistical-learning-for-predictive-modeling-r/"
---

### Load required libraries
```{r}
suppressWarnings({library(omicade4)
library(mogsa)
library(RSpectra)
library(lubridate)
library(dplyr)
library(cowplot)
library(ggplot2)
library(reshape2)
library(DBI)
# library(RPostgreSQL)
library(tidyr)
library(readr)
library(tibble)
library(tidyverse)
library(BatchQC)
library(sva)
library(RcmdrMisc)
library('rstatix')
library("Hmisc")
library(corrplot)
library(GGally)
library(gridExtra)
library(ggpubr)
library(data.table)
library(here)
  })
```


### Set up new directories
```{r}

DEMOGRAPHY_COLS <- c("age_at_boost", "infancy_vac", "biological_sex")

ABTITER_COLS <- c("IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0',
                   "IgG_PT_day14",'IgG_FHA_day14','IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14', 'IgG4_PT_day14','IgG4_FHA_day14')

CELL_COLS <- c("Monocytes_day0", "Monocytes_day1", "CD4Tcells_day0", 
               "CD4Tcells_day3", "ASCs..Plasmablasts._day0", "ASCs..Plasmablasts._day7")

RNA_COLS <- c("CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0", 
              "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14")


"Dataset 2020 ****************************************************************************************************************"
subDir <- "dataset2020"
dataFile <- "dataset2020.csv"
readyToUseFile <- "dataset2020_logTransformed_encoded.csv"

# REMOVE_OUTLIERS <- TRUE
REMOVE_OUTLIERS <- FALSE

LOG_TRANSFORM <- TRUE
# LOG_TRANSFORM <- FALSE

LOG_TRANS_COLS <- c( "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
                    'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')

# NEGATIVE_COLS <-RNA_COLS
NEGATIVE_COLS <- c()

"Dataset 2021 ****************************************************************************************************************"
# subDir <- "dataset2021"
# dataFile <- "dataset2021.csv"
# readyToUseFile <- "dataset2021_logTransformed_encoded.csv"
# 
# # REMOVE_OUTLIERS <- TRUE
# REMOVE_OUTLIERS <- FALSE
# 
# LOG_TRANSFORM <- TRUE
# # LOG_TRANSFORM <- FALSE
# 
# LOG_TRANS_COLS <- c("CCL3_day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0',
#                     'IgG4_PT_day0','IgG4_FHA_day0', "IL6_day0", "NFKBIA_day0", "XIST_day0",
#                     "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
#                     'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')
# 
# # NEGATIVE_COLS <-RNA_COLS
# NEGATIVE_COLS <- c()

"Combined Dataset ****************************************************************************************************************"
# subDir <- "combined_datasets2020_2021"
# dataFile <- "combined_dataset2020_2021_batchCorrected.csv"
# readyToUseFile <- "combined_dataset2020_2021_batchCorrected_logTransformed_encoded.csv"
# 
# # REMOVE_OUTLIERS <- TRUE
# REMOVE_OUTLIERS <- FALSE
# 
# LOG_TRANSFORM <- TRUE
# # LOG_TRANSFORM <- FALSE
# 
# LOG_TRANS_COLS <- c("CCL3_day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0','IgG1_FHA_day0',
#                     'IgG4_PT_day0','IgG4_FHA_day0', "IL6_day0", "NFKBIA_day0", "XIST_day0",
#                     "CCL3_day3", "IL6_day3", "NFKBIA_day7", "XIST_day14", "IgG_PT_day14",'IgG_FHA_day14',
#                     'IgG_PRN_day14','IgG1_PT_day14','IgG1_FHA_day14','IgG4_PT_day14','IgG4_FHA_day14')
# 
# NEGATIVE_COLS <-RNA_COLS
# # NEGATIVE_COLS <- c()

"Create subdirectories ****************************************************************************************************************"
dir.create(here(file.path('./results', subDir)), showWarnings = FALSE)

subDir_trivial <- "1_veryTrivial"
dir.create(here(file.path('./results', subDir, subDir_trivial)), showWarnings = FALSE)

subDir_onePredictor <- "2_onePredictor"
dir.create(here(file.path('./results', subDir, subDir_onePredictor)), showWarnings = FALSE)

subDir_multiPredictor <- "3_multiPredictor"
dir.create(here(file.path('./results', subDir, subDir_multiPredictor)), showWarnings = FALSE)

target_cols <- c("age_at_boost", "Monocytes_day0", "Monocytes_day1","CCL3_day0", "CCL3_day3", "IgG_PT_day0", "IgG_PT_day14")
ranked_cols <- paste("ranked", target_cols, sep="_")

```

### Load Data 
This data was downloaded from CMI PB Database and required features been selected and saved in a CSV files
Some columns been renamed 
Ranked column been generated for age and experimental columns
Categorical features been label encoded
log transformed required columns
```{r}

df_source <- read.csv(here(file.path("./data/", dataFile)))

rankingFunction <- function(x) {
  as.numeric(rank(x, ties.method = "average", na.last = "keep"))
}

df_source[, ranked_cols] <- apply(df_source[, target_cols], 2, rankingFunction)

df_source$infancy_vac <- as.numeric(factor(df_source$infancy_vac, labels = c(1,2), levels = c("wP", "aP"), exclude = NA))
df_source$biological_sex <- as.numeric(factor(df_source$biological_sex, labels = c(1,2), exclude = NA))

df <- df_source

# TO DO: why do we have neg value in Combat results: (I still need to dig into deeper understand the internal process of ComBat but in general it is not considered as good algorithm for RNA seq data "svaseq" considered to be a better choice. Please let me know what do you think)

## "Its probably best to use ComBat on logged data. ComBat assumes that the data symmetric and bell-shaped. If you are still getting negative values, you can set anything negative to 0."

# Combat assumes normally distributed data, which is probably best approximated by log-normalized data. It fits a gene-specific batch effect, which is then subtracted from the data. If some cells in one batch happen to have 0 expression values, this becomes negative. You can deal with these values in two ways:
# 
# Set them to 0.
# Keep the negative values.
# In scenario 1 you interpret the expression values as absolute measurements. 0 is the lower threshold for these. In scenario 2, you interpret the values as relative measurements. Keeping the negative values would conserve the relative expression differences. I wouldn’t take the absolute value though. On another note, setting to 0 also has the added benefit that you may be able to continue using a sparse matrix representation, which is why this is sometimes preferred.

# So the typical analysis workflow would involve the order:
# 
# Read in the data
# Cell and gene QC
# Normalization (+ log transformation if applicable)
# Data correction (includes batch correction and regressing out other covariates).

if (LOG_TRANSFORM){
  for (ltCol in LOG_TRANS_COLS){
    if (ltCol %in% RNA_COLS & length(NEGATIVE_COLS)!=0){
      minCol <- min(df[, ltCol], na.rm = TRUE)
      addedMin <- df[, ltCol]+abs(minCol)+1
      logTf <- log2(df[, ltCol]+abs(minCol)+1)
      df[, ltCol] <- logTf
    }else{
      df[, ltCol] <- log2(df[, ltCol]+1)
    }
  }
}

write.csv(df, here(paste("./results/",subDir, "/", readyToUseFile, sep = "")), row.names = FALSE)
```


### Cumulative distribution function (CDF) and plot
```{r}
cumulCurve <- function(df, colName){
  df <- na.omit(df[, c("subject_id", colName)])
  df <- df[order(df[, colName]),]
  
  df[, 'cumSum'] <- cumsum(df[,2])
  df[, 'orderNumber'] <- 1:dim(df)[1]
  
  ggplot(data=df, aes(x=orderNumber, y=cumSum)) +
  geom_line()+
  geom_point()+
  labs(x="Order Number", y=colName)+
  ggtitle(paste("Outliers of", colName))
}
```


### Plot CDF
```{r}

if (LOG_TRANSFORM){
  expCols <- c(CELL_COLS, RNA_COLS, ABTITER_COLS)
  
  for (i in 1:length(expCols)){
    if (expCols[i] %in% LOG_TRANS_COLS){
      p <- cumulCurve(df_source, expCols[i])
      ggsave(p, file=here(paste("./results/", subDir,"/cumul_notTransformed_", expCols[i],".png", sep="")), width = 5, height = 5 , units = 'in')
      
      p <- cumulCurve(df, expCols[i])
      ggsave(p, file=here(paste("/results/", subDir,"/cumul_transformed_", expCols[i],".png", sep="")), width = 5, height = 5 , units = 'in')
    }else{
      p <- cumulCurve(df_source, expCols[i])
      ggsave(p, file=here(paste("./results/", subDir,"/cumul_", expCols[i],".png", sep="")), width = 5, height = 5 , units = 'in')
    }
  }
}

```


### Box plot function
```{r}
outlierBox <- function(df, varCols, title, dest){
  print(c('varcols', varCols))
  mDf <- melt(df[, varCols])
  print(colnames(mDf))
  
  p<- ggplot(mDf, aes(x=variable, y=value))+
    geom_boxplot(fill="#AE4371")+
    # coord_flip()+
    stat_boxplot(geom='errorbar')+
    xlab(title) +
    theme(axis.title.x = element_text (face = "bold"),
          axis.text.x = element_text(angle = 35, vjust = 0.4, hjust=0.3),
          axis.title.y = element_blank())
  
  ggsave(p, file=dest, width = 2, height = 5 , units = 'in')
}
```


### Outlier detection box plot
```{r}
df <- read.csv(here(paste("./results/", subDir, "/", readyToUseFile, sep="")))

experiments <- list(RNA=c("CCL3_day0", "CCL3_day3"), ab_titer=c("IgG_PT_day0", "IgG_PT_day14"))

for (exp in names(experiments)){
  outlierBox(df, experiments[[exp]], exp, here(paste("./results/", subDir,"/boxPlot_", exp,".png", sep="")))
}
```


### scatter plot and pearson correlation
```{r}
xCols <- c("age_at_boost", "Monocytes_day0", "CD4Tcells_day0", 
           "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0',
           'IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

png(here(paste("./results/", subDir, "/scatterPlot.png", sep = "")), width = 16, height = 16, units = 'in', res = 300 )
ggpairs(data=df[,xCols])
dev.off()

xCols <- c("age_at_boost", "biological_sex", "infancy_vac", "Monocytes_day0", "CD4Tcells_day0", 
           "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0','IgG1_PT_day0',
           'IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

png(here(paste("./results/", subDir, "/scatterPlot_all.png", sep = "")), width = 16, height = 16, units = 'in', res = 300 )
ggpairs(data=df[, xCols])
dev.off()

```

### 1) Very Trivial Analysis 
1. Calculate ranks of predictors e.g. baseline values and age. 
2. Calculate ranks of tasks 
3. calculate spearman correlation of step 1 and 2 ranks
```{r}

x <- c("ranked_age_at_boost", ranked_cols[grepl('day0', ranked_cols)])
y <- ranked_cols[!ranked_cols %in% x ]
df1 <- df

correlation_df<- t(cor(df1[,x], df1[,y], method="spearman", use= "pairwise.complete.obs"))
pcor <- cor.mtest(as.matrix(df1[,ranked_cols]), method="spearman", conf.level = 0.95, na.rm = TRUE)

pcor <- pcor$p[y, x]

colnames(correlation_df) <- gsub("ranked_", "", colnames(correlation_df))
row.names(correlation_df) <- gsub("ranked_", "", row.names(correlation_df))

colnames(pcor) <- gsub("ranked_", "", colnames(pcor))
row.names(pcor) <- gsub("ranked_", "", row.names(pcor))
# 

maxCorr <- max(abs(as.matrix(correlation_df)), na.rm = T)
png(here(paste("./results/", subDir,"/", subDir_trivial, "/correlationHeatmap_spearman_0.95Confidence.png", sep = "")), width = 10, height = 5, units = 'in', res = 300 )
corrplot(correlation_df, tl.col = 'black', addCoef.col = 1,
         p.mat = pcor, sig.level = 0.05, number.cex = 1.5, tl.srt = 45,
         col.lim = c(-maxCorr, maxCorr),
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir,"/", subDir_trivial, "/correlationHeatmap_spearman_0.99Confidence.png", sep="")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(correlation_df, tl.col = 'black', addCoef.col = 1,
         p.mat = pcor, sig.level = 0.01, number.cex = 1.5, tl.srt = 45,
         col.lim = c(-maxCorr, maxCorr),
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```

### Function for regression plot
```{r}
regPlot <- function(df, xCol, yCol,title){
  df$infancy_vac[df$infancy_vac==1] <- "wP"
  df$infancy_vac[df$infancy_vac==2] <- "aP"
  
  p <- ggplot(data = df, aes_string(x= xCol, y = yCol, label= "subject_id")) +
  geom_text(aes(colour=factor(infancy_vac))) +
  scale_color_manual(values=c("red", "springGreen4")) +
  labs(colour= "Vaccine Type") +
  stat_smooth(method = "lm", col = "dodgerblue3") +
  theme(panel.background = element_rect(fill = "white"),
  axis.line.x=element_line(),
  axis.line.y=element_line()) +
  ggtitle(title)+
  stat_cor(method = "spearman")
  }
```


### 2) Linear model with one predictor
1. age been used as predictor for all the tasks
2. the baseline of each task been used as its predictor

## leave one out cross validation been used to estimate model quality
```{r}

set.seed(0)

onePredictor <- list(c("age_at_boost","Monocytes_day1"), c("age_at_boost","CCL3_day3"), c("age_at_boost","IgG_PT_day14"),
                     c("Monocytes_day0", "Monocytes_day1"), c("Monocytes_day0", "CCL3_day3"), c("Monocytes_day0", "IgG_PT_day14"),
                     c("CCL3_day0", "Monocytes_day1"), c("CCL3_day0", "CCL3_day3"),  c("CCL3_day0", "IgG_PT_day14"),  
                     c("IgG_PT_day0", "Monocytes_day1"), c("IgG_PT_day0", "CCL3_day3"), c("IgG_PT_day0", "IgG_PT_day14"))

correlation_pred_true <- data.frame()
pval_pred_true <- data.frame()

for (comb in onePredictor){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", comb)])

  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  
  modelsList <- paste(comb[2], "~", comb[1])
  
  p <- regPlot(subDf, comb[1], comb[2], paste(comb[2], "against", comb[1]))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_onePredictor,"/", comb[2],'_', comb[1], ".png", sep="")), width = 5, height = 5 , units = 'in')
  
  for (i in 1:nrow(subDf)){
    
    fit_1 <- lm(modelsList, data=subDf[-c(i), ])
    
    predictor <- data.frame(subDf[i, comb[1]])
    
    colnames(predictor) <- comb[1]
    predVals <- c(predVals, predict(fit_1, predictor))
    trueVals <- c(trueVals, subDf[i, comb[2]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }

  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(comb[2], "against", comb[1]))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_onePredictor,"/predictedTrue_", comb[1],'_', comb[2], ".png", sep="")), width = 5, height = 5 , units = 'in')
  
  correlations <- cor.test(trueVals, predVals, method="spearman")
  pval_pred_true[comb[1], comb[2]] <- correlations$p.value
  correlation_pred_true[comb[1], comb[2]] <- correlations$estimate
}

maxCorr <- max(abs(as.matrix(correlation_pred_true)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_onePredictor,"/correlationHeatmap_pred_true_spearman_0.95Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(correlation_pred_true)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, na.label = " ",
         p.mat = t(as.matrix(pval_pred_true)), sig.level = 0.05, number.cex = 1.5, tl.srt = 45,
         col.lim = c(-maxCorr, maxCorr),
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_onePredictor,"/correlationHeatmap_pred_true_spearman_0.99Confidence.png", sep = "")), width = 10, height = 5, units = 'in', res = 300 )
corrplot(t(as.matrix(correlation_pred_true)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, na.label = " ",
         p.mat = t(as.matrix(pval_pred_true)), sig.level = 0.01, number.cex = 1.5, tl.srt = 45,
         col.lim = c(-maxCorr, maxCorr),
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```

## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

gList <- list()
coefDf <- data.frame()
pvalDf <- data.frame()

for (comb in onePredictor){
  subDf <- na.omit(df[, comb])
  modelsList <- paste(comb[2], "~", comb[1])
  fit_1 <- lm(modelsList, data=subDf, na.action=na.exclude)
  sum1 <- summary(fit_1)
  mCoeff <- as.matrix(sum1$coefficients)
  coefDf[comb[1], comb[2]] <- mCoeff[comb[1], "Estimate"]
  pvalDf[comb[1], comb[2]] <- mCoeff[comb[1], "Pr(>|t|)"]
}

maxCoef <- max(abs(as.matrix(coefDf)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_onePredictor,"/correlationHeatmap_regression_0.95Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf)), sig.level = 0.05, number.cex = 0.8, na.label = " ",
         col.lim =c(-maxCoef, maxCoef), tl.srt = 45, 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_onePredictor,"/correlationHeatmap_regression_0.99Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf)), sig.level = 0.01, number.cex = 0.8, na.label = " ",
         col.lim =c(min(as.matrix(coefDf), na.rm = TRUE), max(as.matrix(coefDf), na.rm = TRUE)), tl.srt = 45, 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()
```

### 3) Linear model with multiple predictors
Baseline of all three tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

1. only demographic
2. only 3 baseline 
3. 3 baseline + demographic
4. 14 baseline
5. 14 baseline + demographic

### 1. only demographic
## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

correlation_pred_true <- data.frame(row.names = y)
pval_pred_true <- data.frame(row.names = y)

coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)

for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  
  modelsList <- paste(y[j], "~", str_flatten(x, ' + '))
  
  for (i in 1:nrow(subDf)){
    
    fit_1 <- lm(modelsList, data=subDf[-c(i), ])
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    predVals <- c(predVals, predict(fit_1, predictor))
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  
  correlations <- cor.test(trueVals, predVals, method="spearman")
  correlation_pred_true[y[j],"Demographic"] <- correlations$estimate
  pval_pred_true[y[j],"Demographic"] <- correlations$p.value
  
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography"))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_multiPredictor,'/predictedTrue_demographic_', y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}

```

## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex")
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)
for (i in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x, y[i])])
  
  modelsList <- paste(y[i], "~", str_flatten(x, ' + '))
  
  fit_1 <- lm(modelsList, data=subDf[-c(i), ])
  
  sum1 <- summary(fit_1)
  mCoeff <- as.matrix(sum1$coefficients)
  
  coefDf_multi[y[i]] <- mCoeff[2:nrow(mCoeff), "Estimate"]
  pvalDf_multi[y[i]] <-  mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]
}

maxCoef <- max(abs(as.matrix(coefDf_multi)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_demography_0.95Confidence.png", sep = "")), width = 5, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE, 
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.05, number.cex = 0.8,
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_demography_0.99Confidence.png", sep = "")), width = 5, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.01, number.cex = 0.8, 
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()
```
### 2. only baseline
## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CCL3_day0", "IgG_PT_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  
  modelsList <- paste(y[j], "~", str_flatten(x, ' + '))
  
  for (i in 1:nrow(subDf)){
    fit_1 <- lm(modelsList, data=subDf[-c(i), ])
    
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    predVals <- c(predVals, predict(fit_1, predictor))
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  correlations <- cor.test(trueVals, predVals, method="spearman")
  pval_pred_true[y[j], "3 Tasks Baselines"] <- correlations$p.value
  correlation_pred_true[y[j], "3 Tasks Baselines"] <- correlations$estimate

  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "3 Baseline"))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_multiPredictor,"/predictedTrue_3baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}
```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CCL3_day0", "IgG_PT_day0")
                  
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)
for (i in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x, y[i])])
  modelsList <- paste(y[i], "~", str_flatten(x, ' + '))
  
  fit_1 <- lm(modelsList, data=subDf[-c(i), ])
  sum1 <- summary(fit_1)
  mCoeff <- as.matrix(sum1$coefficients)

  coefDf_multi[y[i]] <- mCoeff[2:nrow(mCoeff), "Estimate"]
  pvalDf_multi[y[i]] <-  mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]
}

maxCoef <- max(abs(as.matrix(coefDf_multi)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_3baseline_0.95Confidence.png", sep = "")), width = 5, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.05, number.cex = 0.8,
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_3baseline_0.99Confidence.png", sep = "")), width = 5, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.01, number.cex = 0.8,
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()
```
### 3. Baseline of all three tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex", "Monocytes_day0", "IgG_PT_day0", "CCL3_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])

  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  modelsList <- paste(y[j], "~", str_flatten(x, ' + '))
     
  for (i in 1:nrow(subDf)){
    fit_1 <- lm(modelsList, data=subDf[-c(i), ])
    
    predictor <- data.frame(subDf[i, x])
    colnames(predictor) <- x
    
    predVals <- c(predVals, predict(fit_1, predictor))
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  
  correlations <- cor.test(trueVals, predVals, method="spearman")
  pval_pred_true[y[j], "Demographic + 3 Baselines"] <- correlations$p.value
  correlation_pred_true[y[j], "Demographic + 3 Baselines"] <- correlations$estimate
  # # 
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  # # 
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography + 3 Baseline"))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_multiPredictor,"/predictedTrue_demography_3baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}
```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "biological_sex", "infancy_vac", "Monocytes_day0", "IgG_PT_day0", "CCL3_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)
for (i in 1:length(y)){
  
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x, y[i])])
  modelsList <- paste(y[i], "~", str_flatten(x, ' + '))
  
  fit_1 <- lm(modelsList, data=subDf[-c(i), ])
  sum1 <- summary(fit_1)
  mCoeff <- as.matrix(sum1$coefficients)

  coefDf_multi[y[i]] <- mCoeff[2:nrow(mCoeff), "Estimate"]
  pvalDf_multi[y[i]] <-  mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]
}

maxCoef <- max(abs(as.matrix(coefDf_multi)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_3baseline_demography_0.95Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.05, number.cex = 0.8,
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_3baseline_demography_0.99Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.01, number.cex = 0.8,
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()
```

### 4. 14 baselines
## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0',
       'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  
  modelsList <- paste(y[j], "~", str_flatten(x, ' + '))
  
  for (i in 1:nrow(subDf)){
    fit_1 <- lm(modelsList, data=subDf[-c(i), ])
    predictor <- data.frame(subDf[i, x])

    colnames(predictor) <- x
    
    predVals <- c(predVals, predict(fit_1, predictor))
    trueVals <- c(trueVals, subDf[i, y[j]])
    subjId <- c(subjId, subDf[i, "subject_id"])
    vacType <- c(vacType, subDf[i, "infancy_vac"])
  }
  
  trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
  predVals <- rank(predVals, ties.method = "average", na.last = "keep")
  
  correlations <- cor.test(trueVals, predVals, method="spearman")
  pval_pred_true[y[j], "14 Tasks Baselines"] <- correlations$p.value
  correlation_pred_true[y[j], "14 Tasks Baselines"] <- correlations$estimate
  # # 
  predDf <- data.frame(trueVals, predVals)
  colnames(predDf) <- c('True_values', 'Predicted_values')
  predDf[, "subject_id"] <- subjId
  predDf[, "infancy_vac"] <- vacType
  # # 
  p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against 14 Baseline"))
  ggsave(p, file=here(paste("./results/", subDir, "/", subDir_multiPredictor,"/predictedTrue_14baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
}

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_pred_true_spearman_0.95Confidence.png", sep = "")), width = 15, height = 10 , units = 'in', res = 300)
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.05, number.cex = 0.8, tl.srt = 45, na.label=" ", 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_pred_true_spearman_0.99Confidence.png", sep = "")), width = 15, height = 10 , units = 'in', res = 300)
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.01, number.cex = 0.8, tl.srt = 45, na.label=" ", 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```


## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0',
       'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")
                   
y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")


coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)
for (i in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x, y[i])])
  print(dim(subDf))
  modelsList <- paste(y[i], "~", str_flatten(x, ' + '))
  
  fit_1 <- lm(modelsList, data=subDf[-c(i), ])
  
  sum1 <- summary(fit_1)
  mCoeff <- as.matrix(sum1$coefficients)

  if (!all(is.na(mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]))){
    coefDf_multi[y[i]] <- mCoeff[2:nrow(mCoeff), "Estimate"]
    pvalDf_multi[y[i]] <-  mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]
  }else{
    coefDf_multi[y[i]] <- NaN
    pvalDf_multi[y[i]] <-  NaN
  }
}

maxCoef <- max(abs(as.matrix(coefDf_multi)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_14baseline_0.95Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.05, number.cex = 1, na.label=" ", 
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_14baseline_0.99Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.01, number.cex = 1, na.label=" ", 
         col.lim =c(-maxCoef, maxCoef),
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()
```


### 5. Baseline of all 14 tasks combined with age, biological sex and infancy vaccine type been used to predict each of the three tasks

## leave one out cross validation been used to estimate model quality
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex", "Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", 
       "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0', 'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 
       'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14")

for (j in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x,y[j])])
  predVals <- c()
  trueVals<- c()
  subjId <- c()
  vacType <- c()
  
  modelsList <- paste(y[j], "~", str_flatten(x, ' + '))
  
  sampNumb <- dim(subDf)[1]
  featureNumb <- dim(subDf)[2]
  
  if (sampNumb<=featureNumb){
    pval_pred_true[y[j], "Demographic + 14 Baselines"] <- NA
    correlation_pred_true[y[j], "Demographic + 14 Baselines"] <- NA
  }else{
    
    for (i in 1:nrow(subDf)){
      
      fit_1 <- lm(modelsList, data=subDf[-c(i), ])
      predictor <- data.frame(subDf[i, x])
      colnames(predictor) <- x
      
      predVals <- c(predVals, predict(fit_1, predictor))
      trueVals <- c(trueVals, subDf[i, y[j]])
      subjId <- c(subjId, subDf[i, "subject_id"])
      vacType <- c(vacType, subDf[i, "infancy_vac"])
    }
    
    trueVals <- rank(trueVals, ties.method = "average", na.last = "keep")
    predVals <- rank(predVals, ties.method = "average", na.last = "keep")
    
    correlations <- cor.test(trueVals, predVals, method="spearman")
    pval_pred_true[y[j], "Demographic + 14 Baselines"] <- correlations$p.value
    correlation_pred_true[y[j], "Demographic + 14 Baselines"] <- correlations$estimate
    # # 
    predDf <- data.frame(trueVals, predVals)
    colnames(predDf) <- c('True_values', 'Predicted_values')
    predDf[, "subject_id"] <- subjId
    predDf[, "infancy_vac"] <- vacType
    # # 
    p <- regPlot(predDf, 'True_values', 'Predicted_values', paste(y[j], "against Demography + 14 Baseline"))
    ggsave(p, file=here(paste("./results/", subDir, "/", subDir_multiPredictor,"/predictedTrue_demography_14baseline_", y[j], ".png", sep="")), width = 5, height = 5 , units = 'in')
  }
}

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_pred_true_spearman_0.95Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300)
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.05, number.cex = 1.5, tl.srt = 45, na.label=" ", 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_pred_true_spearman_0.99Confidence.png", sep = "")), width = 10, height = 5 , units = 'in', res = 300)
corrplot(as.matrix(correlation_pred_true), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = as.matrix(pval_pred_true), sig.level = 0.01, number.cex = 1.5, tl.srt = 45, na.label=" ", 
         cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

write.csv(pval_pred_true, here(paste("./results/", subDir, "/", subDir_multiPredictor,"/p_value_pred_true_spearman.csv", sep = "")))
write.csv(correlation_pred_true, here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlation_pred_true_spearman.csv", sep = "")))

```

## the coefficient of regression line of task agains predictors been estimated for feature selection
```{r}
set.seed(0)

x <- c("age_at_boost", "infancy_vac", "biological_sex", "Monocytes_day0", "CD4Tcells_day0", "ASCs..Plasmablasts._day0", 
       "IgG_PT_day0",'IgG_FHA_day0','IgG_PRN_day0', 'IgG1_PT_day0','IgG1_FHA_day0','IgG4_PT_day0', 
       'IgG4_FHA_day0', "CCL3_day0", "IL6_day0", "NFKBIA_day0", "XIST_day0")

y <- c("Monocytes_day1", "CCL3_day3", "IgG_PT_day14") 

coefDf_multi <- data.frame(row.names = x)
pvalDf_multi <- data.frame(row.names = x)
for (i in 1:length(y)){
  subDf <- na.omit(df[, c("subject_id", "infancy_vac", x, y[i])])
  
  if (dim(na.omit(subDf))[1]>length(x)){
    
    modelsList <- paste(y[i], "~", str_flatten(x, ' + '))
    fit_1 <- lm(modelsList, data=subDf)
    sum1 <- summary(fit_1)
    mCoeff <- as.matrix(sum1$coefficients)
    
    coefDf_multi[y[i]] <- mCoeff[2:nrow(mCoeff), "Estimate"]
    pvalDf_multi[y[i]] <-  mCoeff[2:nrow(mCoeff), "Pr(>|t|)"]
  }else{
    coefDf_multi[y[i]] <- NaN
    pvalDf_multi[y[i]] <-  NaN
  }
}

maxCoef <- max(abs(as.matrix(coefDf_multi)), na.rm = T)
png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_14baseline_demography_0.95Confidence.png", sep = "")), width = 15, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.05, number.cex = 1, na.label = " ",
         col.lim =c(-maxCoef, maxCoef), 
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

png(here(paste("./results/", subDir, "/", subDir_multiPredictor,"/correlationHeatmap_regression_14baseline_demography_0.99Confidence.png", sep = "")), width = 15, height = 5 , units = 'in', res = 300 )
corrplot(t(as.matrix(coefDf_multi)), tl.col = 'black', addCoef.col = 1, is.corr = FALSE,
         p.mat = t(as.matrix(pvalDf_multi)), sig.level = 0.01, number.cex = 1, na.label = " ",
         col.lim =c(-maxCoef, maxCoef), 
         tl.srt = 45, cl.ratio = 0.2, col = colorRampPalette(c("blue", "white","red"))(100))
dev.off()

```

